---
title: "incinerators"
author: "paul bradshaw"
date: "2024-06-19"
output: html_document
---

# Analysing data on incinerators

Data has been downloaded from https://wastedataflow.org/reports/default.aspx with the following selections:

* Authority: England
* Periods: Apr 22 - Mar 23
* Questions: Q100

## Install packages

We are going to use [the `dtrackr` package](https://terminological.github.io/dtrackr/) to "creating a step-by-step summary of the important parts of the wrangling as it actually happened to the dataframe, right into dataframe metadata itself."

```{r packages}
#In general use dtrackr is expected to be installed alongside the tidyverse set of packages. It is recommended to install tidyverse first.
library(tidyverse)
#install dtrackr
install.packages("dtrackr")
#this is already installed but needs updating
install.packages("rlang")
#activate dtrackr 
library(dtrackr)
library(readxl)
```


## Import data

The downloaded data is in the same folder as this notebook.

```{r import data}
#skip the first row, headings are on row 2
df <- read_excel("WasteDataFlow Question Data.xlsx", 
                 skip = 1)
```

We also clean the postcodes so they are all upper case

```{r clean postcodes}
df$FacilityPostCode <- str_to_upper(df$FacilityPostCode)
```


## Filter to incineration

```{r filter}

df_filtered <- df %>% 
  track(
    .headline = "WasteDataFlow:",
    .messages = c(
      "loaded from https://wastedataflow.org/reports/default.aspx",
      "skip first row",
      "convert postcodes to upper case",
      "starts with {.count} items")
  ) %>%
  status(
    .messages = "filter by FacilityType == Incineration with/out energy recovery"
  ) %>%
  filter(FacilityType == "Incineration with energy recovery" | FacilityType == "Incineration without energy recovery") %>%
  status(
    .messages = "now what?"
  )

#check the results - should only be two categories
table(df_filtered$FacilityType)
```

Let's see the flowchart so far.

```{r show as flowchart}
df_filtered %>% flowchart()
```

## Group by postcode


```{r group by postcode}
df_grouped <- df %>% 
  track(
    .headline = "WasteDataFlow:",
    .messages = c(
      "loaded from https://wastedataflow.org/reports/default.aspx",
      "skip first row",
      "convert postcodes to upper case",
      "starts with {.count} items")
  ) %>%
  status(
    .messages = "filter by FacilityType == Incineration with/out energy recovery"
  ) %>%
  filter(FacilityType == "Incineration with energy recovery" | FacilityType == "Incineration without energy recovery") %>%
  status(
    .messages = c("Grouped by postcode") #used a vector here as had 2 strings before
  ) %>%
  status(
    .messages = "Summarised by sum of TotalTonnes"
  ) %>%
  group_by(FacilityPostCode) %>%
  summarise(tonnesSUM = sum(TotalTonnes)
            ) %>%
  status(
    .messages = "Remove spaces from postcodes"
  ) %>%
  mutate(Postcode = str_remove_all(FacilityPostCode, "\\s"))

#show flowchart
df_grouped %>% flowchart()
```


## Fetch deprivation for each postcode

Postcode to LSOA lookup for 2011 is downloaded from https://geoportal.statistics.gov.uk/datasets/730eab4ed76b4681ba74317f7e1f1405/about 

Indices of multiple deprivation is https://www.gov.uk/government/statistics/english-indices-of-deprivation-2019 

```{r import lsoa lookup and imd data}
lookupcsv = "Postcode_to_Output_Area_to_Lower_Layer_Super_Output_Area_to_Middle_Layer_Super_Output_Area_to_Local_Authority_District_November_2018_Lookup_in_the_UK.csv"
lookupdf <- read_csv(lookupcsv)

imdxls = "File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx"
imd_df <- read_excel(imdxls,
                     sheet = 2)
```


### Remove spaces from postcodes in each

Some postcodes only match once spaces are removed

```{r remove spaces from postcodes}
#replace spaces and create a new column in the incinerator df
df_grouped$postcode_nospace <- stringr::str_replace_all(df_grouped$FacilityPostCode, " ", "")
#replace spaces and create a new column in the lookup df
lookupdf$postcode_nospace <- stringr::str_replace_all(lookupdf$pcd7, " ","")
```
## Join with LSOA lookup


```{r left join lsoa}

df_joined <- left_join(df_grouped, 
                       lookupdf, 
                       by = "postcode_nospace")
```

```{r select cols}
df_select <- df_joined %>%
  select(colnames(df_joined)[c(1:4,12,15)])

df_select

```

```{r left join IMD}
#create a duplicate column with the same name
imd_df$lsoa11cd <- imd_df$`LSOA code (2011)`

df_joined_imd <- left_join(df_select, 
                       imd_df, 
                       by = "lsoa11cd")

print(colnames(df_joined_imd))

df_joined_imd <- df_joined_imd %>% 
  select(
    colnames(df_joined_imd)[c(1:3,6,7,12)]
  )

df_joined_imd
```

```{r group by IMD}
groupby_imd <- df_joined_imd %>%
  group_by(`Index of Multiple Deprivation (IMD) Decile`) %>%
  summarise(sum(tonnesSUM))

groupby_imd 
```

## Export

```{r export CSV}
write.csv(groupby_imd, "groupby_imd.csv")
write.csv(df_joined_imd, "df_joined_imd.csv")
```


## Show flow chart

Let's try to add those steps for the flowchart.

```{r for flowchart}
df_grouped <- df %>% 
  track(
    .headline = "WasteDataFlow:",
    .messages = c(
      "loaded from https://wastedataflow.org/reports/default.aspx",
      "skip first row",
      "convert postcodes to upper case",
      "starts with {.count} items")
  ) %>%
  status(
    .messages = "filter by FacilityType == Incineration with/out energy recovery"
  ) %>%
  filter(FacilityType == "Incineration with energy recovery" | FacilityType == "Incineration without energy recovery") %>%
  status(
    .messages = c("Grouped by postcode",
                  "{.count} postcodes/rows") 
  ) %>%
  status(
    .messages = "Summarised by sum of TotalTonnes"
  ) %>%
  status(
    .messages = "Clean postcodes (remove spaces)"
  ) %>%
  status(
    .messages = "Look up 2011 LSOA for postcode"
  ) %>%
  status(
    .messages = "Look up deprivation decile for LSOA"
  ) %>%
  group_by(FacilityPostCode)  %>%
  status(
    .messages = "CAN YOU SEE ME 1?"
  ) %>%
  summarise(tonnesSUM = sum(TotalTonnes)
            ) %>%
  status(
    .messages = "CAN YOU SEE ME 2?"
  ) %>%
  mutate(postcode_nospace = str_remove_all(FacilityPostCode, "\\s")) %>%
  status(
    .comment = "CYSM3"
  ) %>%
  left_join(lookupdf, 
                       by = "postcode_nospace") %>% 
  status(
    .messages = "CAN YOU SEE ME 4?"
  ) %>%
  #make sure imd_df has a column with the same name
  left_join(imd_df, 
                       by = "lsoa11cd") %>%
  status(
    .messages = "CAN YOU SEE ME 5?"
  ) %>%
  group_by(`Index of Multiple Deprivation (IMD) Decile`) %>% ##seems to automatically generate a flow chart element
  summarise(sum(tonnesSUM))  %>%
  status(
    .messages = "Summarised by sum of TotalTonnes"
  )

#show flowchart
df_grouped %>% flowchart()
```